---
title: "Assessing the risk of child labor in Ethiopia"

header-includes:
   - \usepackage{array}
date: ""
output: 
  pdf_document:
    fig_height: 3
    fig_width: 5
  html_document:
    fig_height: 3
    fig_width: 5
  word_document:
    fig_height: 3
    fig_width: 5
---

```{r, setup, include = FALSE}
require(MASS)
require(mosaic)
require(dplyr)
require(tidyr)
require(ggplot2)
require(xtable)
require(lme4)
require(gridExtra)
select <- dplyr::select

# Some customization.  You can alter or delete as desired (if you know what you are doing).
trellis.par.set(theme=theme.mosaic()) # change default color scheme for lattice
knitr::opts_chunk$set(
  tidy=FALSE,     # display code as typed
  size="small")   # slightly smaller font for code
options(xtable.comment = FALSE)
```

```{r, include = FALSE}
# Loading in all datasets and necessary files

# Loading data for counts for child labor questions

source("countsForQuestions.R")
source("ReportModelsFinalEthiopia.R")

# Loading in Ethiopia datasets

ethiopia <- read.csv("../Data/FinalData/ethiopiaFinal.csv",
                     stringsAsFactors = FALSE)
ethiopiaHousehold <- read.csv("../Data/FinalData/ethiopiaHousehold.csv",
                     stringsAsFactors = FALSE)
ethiopiaIndividual <- read.csv("../Data/FinalData/ethiopiaIndividual.csv",
                     stringsAsFactors = FALSE)


```

# 4. Results

## 4.2 Exploratory Data Analysis of Outcome Variables

In this section, the distribution for each outcome variable is assessed, as well as the relationships between predictor and outcome variables. All variables in this section are assessed through logistic regression modeling. 

### 4.2.1 Ethiopia Household Level EDA

```{r, echo = FALSE, fig.width = 7, fig.height = 4}
# Counts for hazard ethiopia household

p1 <- ggplot(ethiopiaHouseholdm, aes(hazardCat)) + geom_bar(fill = "aquamarine3") +
  xlab("Are there any children in hazardous work?") + ylab("Number of households") +
  ggtitle("Ethiopia households with \nand without hazardous child labor") +
  geom_text(stat='count', aes(label=..count..), vjust=-.5) +
  scale_y_continuous(limits = c(0, 1550))
  
# Counts for worst ethiopia household

p2 <- ggplot(ethiopiaHouseholdm, aes(worstCat)) + geom_bar(fill = "aquamarine3") +
  xlab("Are there any children in worst child labor?") + ylab("Number of households") +
  ggtitle("Ethiopia households with \nand without worst child labor") +
  geom_text(stat='count', aes(label=..count..), vjust=-.5) +
  scale_y_continuous(limits = c(0, 1250))
grid.arrange(p1, p2, ncol = 2)
```

```{r, echo = FALSE, results = 'asis'}
# hazard and worst counts for Ethiopia Households that overlap

hazardWorstEH <- ethiopiaHouseholdm %>%
  group_by(hazardCat, worstCat) %>%
  summarize(n = n()) %>%
  ungroup() %>%
  spread(hazardCat, n) %>%
  arrange(desc(Yes)) %>%
  select("Any children in worst child labor?" = worstCat,
         "Hazardous Labor: Yes" = Yes, 
         "Hazardous Labor: No" = No) %>%
  as.data.frame()
  
print(xtable(hazardWorstEH, caption = 'Overlap between worst and hazardous labor outcomes for Ethiopia households'),
      include.rownames = FALSE, caption.placement = "top")
```


### 4.4.1 Female vs. Male Headed Households

```{r, echo = FALSE}
# Code needed to make graphs and tables in this section

# number of households per head type in Ethiopia

eTextHead <- ethiopiaHouseholdm %>%
  select(headType) %>%
  group_by(headType) %>%
  summarize(n = n()) %>%
  ungroup() %>%
  spread(headType, n)

# Labels for Graphs
eHeadLabels <- c(`Adult Male` = "Adult Male\n(n = 1272)",
                 `Married Adult Female` = "Married Adult Female\n(n = 250)",
                 `Single Adult Female` = "Single Adult Female\n(n = 283)")

# Ethiopia Household Data for Graphs (Percentage of Houses with Child in Hazardous Child Labor)

eHouseHeadHazard <- ethiopiaHouseholdm %>%
  mutate(hazardCat = ifelse(childNumDefiniteHazardCat == 1, "Yes", "No"),
         worstCat = ifelse(childNumDefiniteWorstCat == 1, "Yes", "No")) %>%
  select(headType, hazardCat) %>%
  group_by(headType, hazardCat) %>%
  summarize(n = n()) %>%
  mutate(total = ifelse(headType == "Adult Male", 1272, 
                        ifelse(headType == "Married Adult Female", 250, 283)),
         percent = round((n/total)*100, 1),
         percentLabel = paste0(percent, "%", sep = ""))
         
# Ethiopia Household Data for Graphs (Percentage of Houses with Child in Worst Child Labor)

eHouseHeadWorst <- ethiopiaHouseholdm %>%
  mutate(hazardCat = ifelse(childNumDefiniteHazardCat == 1, "Yes", "No"),
         worstCat = ifelse(childNumDefiniteWorstCat == 1, "Yes", "No")) %>%
  select(headType, worstCat) %>%
  group_by(headType, worstCat) %>%
  summarize(n = n()) %>%
  mutate(total = ifelse(headType == "Adult Male", 1272, 
                        ifelse(headType == "Married Adult Female", 250, 283)),
         percent = round((n/total)*100, 1),
         percentLabel = paste0(percent, "%", sep = ""))

```

The figure below displays the number of children in hazardous labor per household in Ethiopia, separated by household head. Note instead of a binary indicator of whether or not the house has any children in hazardous work, the exact number of children in hazardous work in the household are shown. The number of observations for each type of household head is displayed as "n", and instead of counts, the instances of how many houses contain a certain number of children in hazardous labor are displayed as proportions. For example, about 50% of single adult female headed houses have exactly one child in hazardous labor. Looking at these proportions across the three groups, adult male headed houses tend to have no children in hazardous labor at a higher rate than both married and single adult female headed houses. 

In fact, 25% of adult male headed households have no children in hazardous labor, meaning that 75% of adult male headed houses have at least one child in hazardous labor. Approximately 80% married adult female headed houses have at least one child in hazardous labor, and the same is true for single adult female headed houses. Among the houses that do have a child in hazardous labor, all household types tend to have only one child in hazardous labor, but adult male headed households proportionally have more households with more than one child in hazardous labor.

When this variable was included in predicting hazardous labor for Ethiopia households, it was found that both married adult female heads and single adult female heads were significant predictors, and associated with higher odds of hazardous labor compared to male headed households. Therefore, there is evidence to suggest that male headed households are associated with lower odds of housing a child in hazardous labor.


```{r, echo = FALSE, fig.height = 4, fig.width = 7}

# plot for Percentage of Houses with Child in Hazardous Child Labor

ggplot(eHouseHeadHazard, aes(hazardCat, percent)) + 
  geom_bar(stat = "identity", fill = "aquamarine3") + 
  facet_wrap(~headType, labeller = as_labeller(eHeadLabels)) +
  xlab("Is there a child (ages 5 to 17) in Hazardous Labor?") +
  ylab("Percentage of Households") + 
  ggtitle("Percentage of Houses with Children in Hazardous Labor \nSeparated by Household Head Type in Ethiopia") +
  geom_text(aes(label=eHouseHeadHazard$percentLabel), vjust=-.5) +
  scale_y_continuous(limits = c(0, 100))
```

The figure below displays the same information except with children in the worst forms of child labor. The majority of all households have at least one child in the worst form of child labor, but adult males proportionally have more households with no children in the worst forms of child labor (with almost 40%) than married and single adult female households, which have 35% and 28% households with no children in the worst forms of child labor, respectively. Therefore, based on this graphic, it appears that female headed households (both married and single) have higher instances of having a child in child labor than adult male headed houses in Ethiopia.

When this parameter was included in predicting worst child labor instances at a household level in Ethiopia, single adult female headed households were significantly associated with increased odds of worst child labor household instances compared to male headed households. However, married adult female headed households were not found to be significantly different than male headed households. 

```{r, echo = FALSE, fig.height = 4, fig.width = 7}

# plot for Percentage of Houses with Child in Worst Child Labor

ggplot(eHouseHeadWorst, aes(worstCat, percent)) + 
  geom_bar(stat = "identity", fill = "aquamarine3") + 
  facet_wrap(~headType, labeller = as_labeller(eHeadLabels)) +
  xlab("Is there a child (ages 5 to 17) in Worst Labor?") +
  ylab("Percentage of Households") + 
  ggtitle("Percentage of Houses with Children in Worst Child Labor \nSeparated by Household Head Type in Ethiopia") +
  geom_text(aes(label=eHouseHeadWorst$percentLabel), vjust=-.5) +
  scale_y_continuous(limits = c(0, 100))
```
